---
title: "Structural Biolinformatics (pt. 1)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Download a CSV file from the PDB site (accessible from “Analyze” > “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and use it to answer the following questions:

Q: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

```{r}
stats <- read.csv("Data Export Summary.csv", row.name=1)
stats
Xray_and_Electron = stats[1,5] + stats[3, 5]
total = sum(stats$Total)
percentage = Xray_and_Electron/total*100
percentage
```

Q: What proportion of structures in the PDB are protein?

```{r}
ans = round(sum(stats$Proteins)/sum(stats$Total) *100, 2)
ans
```


Q: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

1289

## 2. Using bio3d to analyse protein structure data

Use the bio3d package from CRAN.

```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

```{r}
library(bio3d)
aa <- get.seq("1ake_A")
```
```{r}
aa
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')
```

# Download releated PDB files
```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

# Align releated PDBs
```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```

```{r}
pc <- pca(pdbs)
plot(pc)
```

```{r}
library(bio3d.view)
library(rgl)
view(pdbs)
rglwidget(elementId = "tmp")
```


